---
title: Randomization assessment
author: Jake Bowers
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  pdf_document:
    toc: false
---


This file assesses the randomization. We have no reason to doubt the YouGov
randomizer. However, this is our standard practice to check to learn about
possible chance covariate imbalances. Notice that we do not use survey weights
here: our goal is to ask whether the randomization done *among the sampled
respondents* worked as it should --- i.e had no systematic relationship with a
collection of covariates.


```{r setup, echo=FALSE}
library(here)
source(here("analysis","rmd_setup.R"))
library(coin)
library(RItools)
library(estimatr)
library(foreign)
```

```{r readdata, echo=TRUE, warning=FALSE}
dat <- read.spss(here("data/Fourth_Wave","TPL_Testing_Survey_FourthWave_YouGov_MERGEDWITHTHIRDWAVEFORMESSING.sav"),to.data.frame = TRUE)
```

```{r datasetup, echo=TRUE}
## Treatment assignment q115_treat

# Looks like not everyone was included in the
## experiment in the fourth wave, that 500 people were excluded (perhaps not
## included in this wave)

table(dat$q115_treat,exclude=c())

## So, just focus on the valid respondents.

datw4 <- droplevels(dat[!is.na(dat$q115_treat),])

## Some of the code below wants either treatment assignment or outcome to be either a factor variable (with labels) or a binary variable.
table(datw4$q115_treat,exclude=c())
datw4$q115N <- as.numeric(datw4$q115_treat=="Community")
datw4$q115F <- factor(datw4$q115N)
```

We think that treatment assignment was done within strata of pid3 but are not sure. For now, presenting the analysis **both** ways.

Since I am not 100% sure I understand all of the recoding, I present a couple
of different approaches here. First, I try to let R handle the categorical vs
numeric variable issue. I think these are the relevant covariates where a
chance imbalance could change our interpretation of the results.

Here I present the results for **both** the simply or completely randomized
case ("unblocked") and the block-randomized case (where randomization occurred
within strata/blocks of party id) ("pid").


```{r randtest1, echo=TRUE}
## Trying to let R handle the dummy recoding
xb1 <- xBalance(q115N~q1+age+gender_client+educ+dependents_dummy_coded+race+core_city,
    strata=list(notblocked=NULL,pid3=~pid3),data=datw4,report="all")

## Overall no detectable departure from what we'd expect from a well executed random assignment
xb1$overall

## And here just comparing the one-by-one p-values (without adjustment)

theps <- xb1$results[,"p",]
theps_adj <- apply(theps,2,function(x){ p.adjust(x,method="holm") })
pres <- data.frame(cbind(unadj=theps,adj=theps_adj))
names(pres)[3:4] <- paste("Adj:",names(pres)[3:4])
pres

## More information (including unadjusted p-values)
xb1$results[,c("q115N=0","q115N=1","adj.diff","std.diff","p"),]

```

Checking the asymptotic assumption of `xBalance` (compare to `xb1$overall` above)

```{r coincheck, echo=TRUE}

coin1_asympt <- independence_test(q1+age+gender_client+educ+dependents_dummy_coded+race+core_city~q115F|pid3,data=datw4,
    teststat="quadratic",distribution=asymptotic())
coin1_asympt
pvalue(coin1_asympt)

coin1_perm <- independence_test(q1+age+gender_client+educ+dependents_dummy_coded+race+core_city~q115F|pid3,data=datw4,
    teststat="quadratic",distribution=approximate(nresample=5000))
coin1_perm
## This next shows a "confidence interval" because another 5000 sims might change the p-value
pvalue(coin1_perm)

```


Here is a version using the recoding from SPSS. I'm not sure why we have a difference.

```{r}
## Covariates
## q1 age gender_client race educ pid3 faminc_new
## q115_treat age gender_dummy_coded_female party_dummy_coded_republican
##    race_dummy_coded_black race_dummy_coded_latino core_city_excluding_Newport_and_noncentral_PVD
##    eighteen_to_twentynine_dummy_coded sixty_and_over_dummy_coded low_SES_dummy_coded No_HS_dummy_coded
##    HS_dummy_coded Some_college_dummy_coded two_year_degree_dummy_coded dependents_dummy_coded.


xb2<-xBalance(q115N~age+gender_dummy_coded_female+party_dummy_coded_republican+race_dummy_coded_black+
    race_dummy_coded_latino+core_city_excluding_Newport_and_noncentral_PVD+
    eighteen_to_twentynine_dummy_coded+sixty_and_over_dummy_coded+low_SES_dummy_coded+No_HS_dummy_coded+
    HS_dummy_coded+Some_college_dummy_coded+two_year_degree_dummy_coded+dependents_dummy_coded,
strata=list(notblocked=NULL,pid3=~pid3),data=datw4,report="all")

xb2$overall

xb2$results[,c("q115N=0","q115N=1","adj.diff","std.diff","p"),]
xb2$results[,"p",]

```
